---
title: "P curve under no effect"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, comment=NULL)
```

It is important to understand that the $p$ value is itself a random variable.  In other words, the same (statistical) question will give raise to varying results everytime an experiment is carried out to try to answer it.

How consistent an answer the $p$ value can give can be investigated *in silico* using replication and the `replicate()` function.

Let's first load the required libraries:

```{r}
library("tidyverse")
```

Here two sets of numbers $x$ and $y$ are drawn from the same distribution.

```{r exercise1, cache=TRUE}
R = 100000 # Number of replicated
n = 10     # Group size

psim1 = replicate(R, {
    x = rnorm(n, mean=0, sd=1)
    y = rnorm(n, mean=0, sd=1)
    
    t = t.test(x, y)
    t$p.value
})
```

The distribution can be shown using `hist(psim1)` or the ggplot2 functions:

```{r}
qplot(psim1, geom="histogram", breaks=seq(0, 1, len=30))
```

A perhaps more pleasant result can be obtained using

```{r}
ggplot(tibble(p=psim1), aes(p, ..density..)) +
    geom_histogram(breaks=seq(0, 1, len=30),
                   fill='white', colour='gray') +
    ggtitle("Distribution of the p value under the null")
```

This histogram is called a "$p$ curve".

The interesting point is that a $p$ value when the null hypothesis is indeed true is uniformly distributed between $0$ and $1$.  The $p$ value behaves as if it is drawn from a uniform distribution `runif(1)`.  When the null hypothesis is true, the $p$ curve is flat.

Note that the flatness means that occasionally, $p$ values will reach very small values. In fact, $5$% of the time, the $p$ value will reach significance levels of $0.05$ since the distribution is flat.